---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(readr)
library(dplyr)
library(corrplot)
library(GGally)
library(FactoMineR)
library(factoextra)
library(ggplot2)
library(tidyr)
library(MASS)
```

# **4. Dimensionality reduction techniques - week 5**
```{r}
date()
```
# 4.1  _Data overview _   

Description of 'human' dataset variables.      
Original data from: http://hdr.undp.org/en/content/human-development-index-hdi    
Retrieved, modified and analyzed by Tuomo Nieminen 2017    

 The data combines several indicators from most countries in the world
"Country" = Country name     
There are in total 8 variables explaining human development, considered here:    

Related to health and knowledge:   
EDU.EDP: Expected years of schooling, LIFE.EXP: Life expectancy at birth, GNI: The gross national income Index, MAT.MOR: Maternal mortality, ADO.BIRTH: Adolescent birth ratio.   



Related to empowerment:   
EDU2.F: Percentage of females having secondary education, LABO.F: Percentage of females being employed, PARLI.F: Female representation in parliament.  


```{r}
human <- read_csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt")
str(human)
```
Correlation between mentioned variables:
```{r}
human <- read.csv("~/GitHub/IODS-project/IODS-project/data_AKMV/human2.txt")
ggpairs(human)

```

# 4.2 _PCA_ (PRINCIPAL COMPONENT ANALYSIS)
PCA raw data:       
```{r}
human <- read.csv("~/GitHub/IODS-project/IODS-project/data_AKMV/human2.txt")
pca_human<-prcomp(human)
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```
          
      
contribution to the dimensions:

```{r}
res.pca1 <- prcomp(human, center=TRUE, scale = TRUE)
fviz_pca_var(res.pca1, col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE, title="Not-standardized variables")

```

we can see that PC1 explains 53.6% of the systems while PC2 only 16.2%., the variables that better explain the system are: life expectancy, maternal mortality and education expectancy. Similarly in PC2 is the presence of working females.

PCA of standardized variables:
```{r}
human_std <- scale(human)
pca_human <- prcomp(human_std)
biplot(pca_human,repel=TRUE,  choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```
          
      
contribution to the dimensions:

```{r}
res.pca <- prcomp(human_std, center=TRUE, scale = TRUE)
fviz_pca_var(res.pca, col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),repel = TRUE, title="Standardized variables")
```

The relevance of the variables does not changed, but how the countries are distributed does.

# 4.3 _MCA_ (MULTIPLE CORRESPONDENCE ANALYSIS)

The data used here concern a questionnaire on tea. We asked to 300 individuals how they drink tea (18 questions), what are their product's perception (12 questions) and some personal details (4 questions).    
(https://rdrr.io/cran/FactoMineR/man/tea.html)

Overview of the data:    
```{r}
data(tea)
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time3 <- dplyr::select(tea, one_of(keep_columns))

gather(tea_time3) %>% ggplot(aes(value)) + 
  facet_wrap("key", scales = "free") + geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

Making the MCA graphs:
```{r}
mca <- MCA(tea_time3, graph = FALSE)
fviz_mca_var(mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE,
             ggtheme = theme_minimal())

```

Having a look at the variables 
```{r}
fviz_contrib(mca, choice = "var", axes = 1, top = 15)
```


 